Redundant calibration as a complex optimization problem

### 2017-01-12 paper.tex
added dimension table

New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also recently sparked a renewed interest in redundant calibration. Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used to detect the EoR 21 cm signal.

@@

Two ways of speeding up redundant calibration have also been proposed recently. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
the true observed visibilities \citep{Wijnholds2012,Marthi2014}. The second is to use the preconditioned conjugate gradient (PCG) method. PCG allows us to exploit the sparsity inherent in the redundant calibration problem, 
to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.

In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
Using the LM algorithm, reduces the number of algorithmic layers that are needed to perform redundant calibration. 
We also compare two recent redundant calibration speedup algorithms with each other, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it 
allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm. Moreover, this framework also 
allows us to show that the ADI method proposed by \citep{Marthi2014} can for all intents and purposes be renamed to redundant \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}.

ADD SUMMARY PARAGRAPH WHICH DEPENDS ON PAPER LAYOUT

@@

\section{Preliminaries}
There are two major pre-existing algorithms we use throughout this paper, namely the Conjugate Gradient (CG) method and least-squares parameter estimation.
We respectively discuss the CG method and least-squares in Section~\ref{sec:conj_grad} and Section~\ref{sec:ls}.

@@

\begin{algorithm}
\caption{Conjugate Gradient Method. Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}. \citep{Shewchuk1994}.}\label{algo:CG}
\begin{algorithmic}[1]
\State $k \gets 0$
\State $\br \gets \bb - \bA\bx$
\State $\bp \gets \br$
\State $\delta_{\textrm{new}} \gets \br^H\br$
\State $\delta_0 \gets \delta_{\textrm{new}}$
\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
\State $\alpha_k \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
\State $\bx \gets \bx + \alpha_k\bp$
\State $\br \gets \br - \alpha\bA\bp$
\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
\State $\delta_{\textrm{new}} \gets \br^H\br$
\State $\beta_{k} \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
\State $\bp \gets \br + \beta\bp$
\State $k \gets k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

@@

\subsubsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix is equal to one).
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

@@

The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm. In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}. 
The GN update step is defined as
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
\end{equation}
while the LM update takes the following form
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
\end{equation}
where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper we respectively denote 
matrix transposition and matrix inversion with $(*)^T$ and $(*)^{-1}$.
These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.

@@

\section{Problem Statement}
In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
We present the Signal-to-noise ratio (SNR) definition we used in this paper in Section\ref{sec:snr}. The general scope and limitations of this paper is discussed in Section~\ref{sec:scope}.

@@

Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
The aforementioned fact allows us to group all the redundant baselines in a redundant array into the same redundant baseline group. When our array 
is a redundant array we therefore only require single index values for each redundant baseline group, instead of a unique single index label for 
each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
\begin{equation}
\zeta_{pq} = 
\begin{cases}
\phi_{pq}~\textrm{if}~p \leq q\\
\phi_{qp}~\textrm{if}~p>q
\end{cases}.
\end{equation} 
It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
It becomes increasingly more difficult to construct analytic expressions of $\zeta_{pq}$ for other more complicated array layouts. 
We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be construed as the maximal number of unique redundant baseline groups which can be formed for a given 
array layout. Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices. In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry instead. The dimension of these geometry 
matrices are $N\times N$.

@@

If the function $\phi_{pq}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline group index itself then it is possible 
to calculate the unknown antenna index. This can be computed via the following two expressions:
\begin{equation}
\xi_{ij} = 
\begin{cases}
p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
0~\textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
\psi_{ij} = 
\begin{cases}
q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
0~\textrm{otherwise}
\end{cases}
\end{equation}
We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline group index are known, while we use $\psi_{ij}$ to obtain 
a similar result: the second antenna index of the composite antenna pair. %\textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.

@@

\subsection{Signal-to-noise Ratio}
\label{sec:snr}
The following SNR definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

\subsection{Paper Scope}
\label{sec:scope}
We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}. Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.

@@

The aforementioned is achieved by first rewriting Eq.~\eqref{eq:least_squares_red} to
\begin{equation}
\label{eq:least_squares_lincal}
\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
\end{equation}
where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the \textsc{lincal} algorithm (see Appendix~\ref{sec:lincal}).
The Jacobian associated with Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. 
At the hand of Appendix~\ref{sec:lincal} it becomes evident that \textsc{lincal} is equivalent to a basic GN update step \citep{Kurien2016}. 
Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.

@@

\subsubsection{Complex GN and LM}
\label{sec:complex_GN_LM}
In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
where 
\begin{align}
\bmJ &= \frac{\partial \bv}{\partial \bz}, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{align}
In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator \citep{Sorber2012}.
The GN update step is now defined as:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
\end{equation}
The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}
where $\bD=\bI\odot\bJ^H\bJ$. We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$.

@@

\subsubsection{Analytic Expressions}
If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
\begin{equation}
\label{eq:Jacobian_red}
\bJ = \begin{bmatrix}
       \bM & \bN\\
       \conj{\bN} & \conj{\bM}
      \end{bmatrix},
\end{equation}
where
\begin{equation}
\bM =\begin{bmatrix}
      \bO & \bP
     \end{bmatrix},
\end{equation}
and 
\begin{equation}
\bN = \begin{bmatrix}
       \bQ & \bzero
      \end{bmatrix}.
\end{equation}
Moreover,
\begin{equation}
\left [ \bO  \right ]_{\alpha_{pq},j} = \begin{cases}
                                         y_{\phi_{pq}}\conj{g_q} & \textrm{if}~p=j\\
                                         0  & \textrm{otherwise} 
                                        \end{cases},
\end{equation}

@@

\begin{equation}
[\bE]_{ij} = 
\begin{cases}
 \sum_{pq \in \mathcal{PQ}_i} \left | g_p \right |^2 \left | g_q \right |^2  & \textrm{if} ~ i=j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
\begin{equation}
[\bF]_{ij} = 
\begin{cases}
 g_i g_j  \left | y_{\zeta_{ij}} \right |^2  & \textrm{if} ~ i \neq j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
[\boldsymbol{G}]_{ij} = 
\begin{cases}
 g_i y_j  \left | g_{\xi_{ij}} \right |^2  & \textrm{if} ~ \xi_{ij}\neq0\\
 0 & \textrm{otherwise}
\end{cases}.
\end{equation}
Moreover, $(*)^T$ denotes matrix transposition and
\begin{equation}
\mathcal{PQ}_i = \left\{pq\in\mathbb{N}^2|(\phi_{pq} = i) \right\}.
\end{equation}

Furthermore, substituting Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\breve{\br}$ results in
\begin{equation}
\bJ^H\breve{\br} = \begin{bmatrix}
                   \ba \\
                   \bb \\
                   \conj{\ba}\\
                   \conj{\bb}
                   \end{bmatrix},
\end{equation}
where
\begin{align}
\label{eq:ab}
\left [ \ba \right ]_i &= \sum_{k\neq i} g_k \widetilde{y}_{ik}r_{ik},  & \left [ \bb \right ]_i &= \sum_{pq\in\mathcal{PQ}_i}\conj{g}_p g_q r_{pq},
\end{align}
and
\begin{equation}
\widetilde{y}_{ik} = 
\begin{cases}
\conj{y}_{\zeta_{ik}} & \textrm{if}~k > i\\
y_{\zeta_{ik}} & \textrm{otherwise}
\end{cases}.
\end{equation}
Moreover $\ba$ and $\bb$ are both column vectors.

HAVE TO RECHECK THE ABOVE EQUATIONS DIRECTLY FROM THE CODE.
DEFINED ALL OPERATORS AND SYMBOLS?

@@

>>>>

New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also recently sparked a renewed interest in redundant calibration. Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.

@@

Two ways of speeding up redundant calibration have also been proposed recently. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
the true observed visibilities \citep{Wijnholds2012,Marthi2014}. Note that the fact that \citet{Marthi2014} is making use of the ADI method is not immediately clear
from the paper itself. We will elaborate on this later on in the paper (see Section~\ref{sec:results}). The second is to use the preconditioned conjugate gradient (PCG) method. PCG allows us to exploit the sparsity inherent in the redundant calibration problem, 
to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.

In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
Using the LM algorithm, reduces the number of algorithmic layers that are needed to perform redundant calibration. 
We also compare two recent redundant calibration speedup algorithms with each other, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it 
allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.

%Moreover, this framework also 
%allows us to show that the ADI method proposed by \citep{Marthi2014} can for all intents and purposes be renamed to redundant \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}.

We start the paper by presenting some useful preliminaries in Section~\ref{sec:preliminaries}. In Section~\ref{sec:p_statement}, we formulate redundant calibration 
as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec}, we use the complex framework proposed by \citet{Smirnov2015} to derive a numerical
algorithm (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
We end the paper with some final conclusions.

@@

\section{Preliminaries}
\label{sec:preliminaries}
There are two major pre-existing algorithms we use throughout this paper, namely the Conjugate Gradient (CG) method and the least-squares solver.
We respectively discuss the CG method and least-squares in Section~\ref{sec:conj_grad} and Section~\ref{sec:ls}.

@@

\begin{algorithm}
\caption{Conjugate Gradient Method. Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}. \citep{Shewchuk1994}.}\label{algo:CG}
\begin{algorithmic}[1]
\State $k \gets 0$
\State $\br \gets \bb - \bA\bx$
\State $\bp \gets \br$
\State $\delta_{\textrm{new}} \gets \br^H\br$
\State $\delta_0 \gets \delta_{\textrm{new}}$
\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
\State $\alpha \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
\State $\bx \gets \bx + \alpha\bp$
\State $\br \gets \br - \alpha\bA\bp$
\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
\State $\delta_{\textrm{new}} \gets \br^H\br$
\State $\beta \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
\State $\bp \gets \br + \beta\bp$
\State $k \gets k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

@@

\subsubsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). In this paper, we denote matrix inversion with $(*)^{-1}$. The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix is equal to one).
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

@@

The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm. In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}. 
The GN update step is defined as
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
\end{equation}
while the LM update takes the following form
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
\end{equation}
where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
matrix transposition with $(*)^T$.
These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.

@@

\section{Problem Statement}
\label{sec:p_statement}
In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
We present the Signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.

@@

Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group. Assigning the baselines of a redundant array to redundant baseline groups has the advantage that it reduces the number of single index values we require. We now only require single index values for each redundant baseline group, instead of a unique single index label for 
each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
\begin{equation}
\zeta_{pq} = 
\begin{cases}
\phi_{pq}~\textrm{if}~p \leq q\\
\phi_{qp}~\textrm{if}~p>q
\end{cases}.
\end{equation} 
It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
It becomes increasingly more difficult to construct analytic expressions of $\zeta_{pq}$ for other more complicated array layouts. 
We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be construed as the maximal number of unique redundant baseline groups which can be formed for a given 
array layout. Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices. In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry instead. The dimension of these geometry 
matrices are $N\times N$.

@@

If the function $\phi_{pq}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline group index itself then it is possible 
to calculate the unknown antenna index. This can be computed via the following two expressions:
\begin{equation}
\xi_{ij} = 
\begin{cases}
p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
0~\textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
\psi_{ij} = 
\begin{cases}
q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
0~\textrm{otherwise}
\end{cases}.
\end{equation}
We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline group index are known, while we use $\psi_{ij}$ to obtain 
a similar result: the second antenna index of the composite antenna pair. %\textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.

@@

\subsection{Signal-to-noise Ratio}
\label{sec:snr}
The following SNR definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

\subsection{Limitations}
\label{sec:scope}
We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}. Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.

@@

The aforementioned is achieved by first rewriting Eq.~\eqref{eq:least_squares_red} to
\begin{equation}
\label{eq:least_squares_lincal}
\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
\end{equation}
where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the algorithm implemented in \textsc{lincal} (see Appendix~\ref{sec:lincal}).
The Jacobian associated with Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. 
At the hand of Appendix~\ref{sec:lincal} it becomes evident that \textsc{lincal} is equivalent to a basic GN update step \citep{Kurien2016}. 
Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.
In doing so the need to use $\textsc{logcal}$ before $\textsc{lincal}$ becomes superfluous. The use of $\textsc{logcal}$ may, however, still offer 
a computational advantage, as it is a very fast method to obtain a good starting solution, and would reduce the number of iterations needed by the LM algorithm.
Proper experiments would need to be conducted to determine whether it is worth keeping $\textsc{logcal}$ in the algorithmic hierarchy.

@@

\subsubsection{Complex GN and LM}
\label{sec:complex_GN_LM}
The complex Jacobian is defined to be
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
where 
\begin{align}
\bmJ &= \frac{\partial \bv}{\partial \bz}, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{align}
In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator \citep{Sorber2012}.
The GN update step is now defined as:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
\end{equation}
The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}
where $\bD=\bI\odot\bJ^H\bJ$. We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$.

@@

\subsubsection{Analytic Expressions}
\label{sec:analytic}
If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
\begin{equation}
\label{eq:Jacobian_red}
\bJ = \begin{bmatrix}
       \bM & \bN\\
       \conj{\bN} & \conj{\bM}
      \end{bmatrix},
\end{equation}
where
\begin{equation}
\bM =\begin{bmatrix}
      \bO & \bP
     \end{bmatrix},
\end{equation}
and 
\begin{equation}
\bN = \begin{bmatrix}
       \bQ & \bzero
      \end{bmatrix}.
\end{equation}
Moreover,
\begin{equation}
\left [ \bO  \right ]_{\alpha_{pq},j} = \begin{cases}
                                         y_{\phi_{pq}}\conj{g_q} & \textrm{if}~p=j\\
                                         0  & \textrm{otherwise} 
                                        \end{cases},
\end{equation}

@@

\begin{equation}
[\bE]_{ij} = 
\begin{cases}
 \sum_{rs \in \mathcal{RS}_i} \left | g_r \right |^2 \left | g_s \right |^2  & \textrm{if} ~ i=j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
\begin{equation}
[\bF]_{ij} = 
\begin{cases}
 g_i g_j  \left | y_{\zeta_{ij}} \right |^2  & \textrm{if} ~ i \neq j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
[\boldsymbol{G}]_{ij} = 
\begin{cases}
 g_i y_j  \left | g_{\xi_{ij}} \right |^2  & \textrm{if} ~ \xi_{ij}\neq0\\
 0 & \textrm{otherwise}
\end{cases}.
\end{equation}
Moreover, 
\begin{equation}
\mathcal{RS}_i = \left\{rs\in\mathbb{N}^2|(\phi_{rs} = i) \right\}.
\end{equation}

Furthermore, substituting Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\breve{\br}$ results in
\begin{equation}
\bJ^H\breve{\br} = \begin{bmatrix}
                   \ba \\
                   \bb \\
                   \conj{\ba}\\
                   \conj{\bb}
                   \end{bmatrix},
\end{equation}
where
\begin{align}
\label{eq:ab}
\left [ \ba \right ]_i &= \sum_{k\neq i} g_k \widetilde{y}_{ik}r_{ik},  & \left [ \bb \right ]_i &= \sum_{rs\in\mathcal{RS}_i}\conj{g}_r g_s r_{rs},
\end{align}
and
\begin{equation}
\widetilde{y}_{ik} = 
\begin{cases}
\conj{y}_{\zeta_{ik}} & \textrm{if}~k > i\\
y_{\zeta_{ik}} & \textrm{otherwise}
\end{cases}.
\end{equation}
Moreover $\ba$ and $\bb$ are both column vectors. The length of $\ba$ and $\bb$ is $N$ and $L$, respectively.
The dimensions of the matrices we defined in this section are presented in Table~\ref{tab:matrix_dimensions}.

@@

\begin{table}
\centering
\caption{The dimensions of the matrices defined in Section~\ref{sec:analytic}.}
\begin{tabular}{|c c|} 
\hline
Matrix & Dimension\\
\hline
\hline
$\bJ$ & $2B \times P$ \\
$\bM$ & $B \times (N+L)$ \\
$\bN$ & $B \times (N+L)$ \\
$\bO$ & $B \times N$ \\
$\bP$ & $B \times L$ \\
$\bQ$ & $B \times N$ \\
$\bzero$ & $B \times L$ \\
\hline
\hline
$\bH$ & $P\times P$\\
$\bA$ & $(N+L)\times (N+L)$\\
$\bB$ & $(N+L)\times (N+L)$\\
$\bC$ & $N \times N$\\
$\bD$ & $N \times L$\\
$\bE$ & $L \times L$\\
$\bF$ & $N \times N$\\
$\bG$ & $N \times L$\\
$\bzero$ & $L \times L$\\
\hline
\end{tabular}
\label{tab:matrix_dimensions}
\end{table}
